---
title: "Analysis Walkthrough"
editor: visual
format:
  html:
    self-contained: true
    code-fold: true
    code-tools: true
---

# Percentage of the US Population with a Bachelor's Degree and Higher

We'll start with loading our libraries.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tigris)
library(tidycensus)
library(tmap)
library(tmaptools)
library(janitor)
library(writexl)
library(here)
options(tigris_class = "sf")

```

Below are the census variables use for the data analysis.

```{r}
myvars <- c(education_total = "B06009_001",
            education_bachelors = "B06009_005",
            education_gradprofess = "B06009_006")

```

Using the `get_acs()` function in tidycensus, we will be able to see the previously identified variables on the state-level.

```{r}

education_one <- get_acs(geography = "state",
  variables = c(myvars),
  output = "wide",
  geometry = TRUE
  )

```

To clean up the data set, I used the `select()` function to remove the three "margin of error" columns.

```{r}

education_two <- education_one %>%
  select(-ends_with("M"))

names(education_two)


```

Using the `mutate()` function, I created a new column that contains the percentage of people with a bachelor's or higher degree for each state.

```{r}

education_two_mutated <- education_two %>%
  mutate(
    total_above_bach = education_bachelorsE + education_gradprofessE,
    pct_edu_above_bach = total_above_bach/education_totalE * 100
  )

glimpse(education_two_mutated)

```

After removing Alaska, Hawaii, and Puerto Rico, I made a map of the continental U.S. showing the new percentage measure created in the previous code chunk.

```{r}

mutate_edu_two <- subset(education_two_mutated, NAME != "Alaska")

mutate_two_edu_two <- subset(mutate_edu_two, NAME != "Hawaii")

mutate_final_edu_two <- subset(mutate_two_edu_two, NAME != "Puerto Rico")


```

Using the tmap() package, I created the final version of the shaded map for the continental U.S. that shows the percentage of Bachelor's-and-higher populations in each state.

```{r}

tmap_mode(mode = "plot")

tm_shape(mutate_final_edu_two) +
  tm_polygons("pct_edu_above_bach", id = "GEOID")


```
